Sharing statistics for SPARQL Federation optimization, with emphasis on benchmark quality

### 2012-01-10 eswc-2012/phdsymp.tex
stochastic variable
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@53 43ef1190-90fd-4dd7-a769-d0121f02b80e

\documentclass{llncs}

@@

While the basic feature set of the proposed standard can enable users
to create federated queries, it is not of great use as it requires
extensive prior knowledge of both the data to be queried and
performance characteristics of the involved query engines. Without
this knowledge, the overall performance is insufficient for any
practical applications.

@@

>>>>

%\documentclass{llncs}
\documentclass{article}

@@

While the basic feature set of the proposed standard can enable users
to create federated queries, it is not of great use as it requires
extensive prior knowledge of both the data to be queried and
performance characteristics of the involved query engines. Without
this knowledge, the overall performance is insufficient for many
practical applications.

@@

Current common practice in benchmarking SPARQL-enabled systems is to
use or synthesize a certain dataset, then formulate a number of
queries seen as representative of SPARQL use in some way. These
queries are then executed, and some characteristic of performance is
measured, for example the time it takes for the engine to return the
full result. Commonly, this process is repeated a number of times, and
an average response time is found. The performance of different
engines can then be compared based on these averages.

In many cases, this is sufficient. Sometimes, one engine can execute a
query an order of magnitude faster than another, in which case, there
is hardly reasonable doubt which is faster.

When the difference is little, we must consider the possibility that
the measurements are impacted by random noise. Computer systems are
generally very deterministic systems, but there are nevertheless many
such scenarios: The underlying system may be doing other tasks. This
is likely to be particularly important in enterprise environments
where high-load applications are often running on virtual
machines. The engine may use randomized algorithms, or the data may be
of such a nature it triggers worst-case behaviour in situations the
query writer cannot foresee. All this contributes to that we should
consider the performance metric, whatever metric is used, as a
\emph{stochastic variable}.

Doing this opens new methodological possibilities, first and foremost
using well-established statistical hypothesis testing rather than just
comparing averages.
